---
title: "Practical Machine Learning - Assignment"
author: "Daniel Jungen"
date: "27 Januar 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this document we will compare different prediction algorithms in order to find the one that makes the best predictions under the given assumptions.

The background is stated in the descpription of the assignment:

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

The data we use in that assignment can be downloaded here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

From now on we assume these files to be downloaded to the working directory.


```{r}
training <- read.csv("pml-training.csv")
testing  <- read.csv("pml-testing.csv")
```

#General Preparation

In order to work with the data efficiently we ne to load some packages which we assume to be installed:

```{r message = FALSE}
library(knitr)
library(caret)
library(rpart)
library(rpart.plot)
library(rattle)
library(randomForest)
library(corrplot)
```

An inspection of the classes in the training set tells us that there are five levels to predict:

```{r}
unique(training$classe)
```

The first four columns of the training set can be omitted as they are only meant for identifcation (of the users) but not for any classification purposes. In the same steo we can partition the training set into a smaller training set and a testing set:


```{r}

train_index<-createDataPartition(training$classe, p=.75, list = FALSE)

train_small<-training[train_index,-(1:4)]
test_small<-training[-train_index,-(1:4)]
```

In order to get a training set to build reliable predictions on the training set we should  remove those predictors (columns) that have a very low variance or a lot of values which are (almost) zero:

```{r}
nearzv<-nearZeroVar(train_small)

train_small<-train_small[,-nearzv]
test_small<-test_small[,-nearzv]


almostna    <- sapply(train_small, function(x) mean(is.na(x))) > 0.95
train_small <- train_small[, almostna==FALSE]
test_small  <- test_small[, almostna==FALSE]
```

So now the 
